---
title: "Classwork"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In class work from 2017-10-16.

# Poisson overdispersion: seedling data

## Regular Poisson

```{r}
seed <- read.csv("../data/Brsp_seedl2.csv", stringsAsFactors = F)

seedl.pois <- glm(Recruits ~ Basal, 
                  family = poisson,
                  data = seed)

summary(seedl.pois)

deviance(seedl.pois)/df.residual(seedl.pois)
86.74/16 # resid dev : df is > 1
```

Based on the ratio of residual deviance to degrees of freedom, this looks overdispersed.

## Quasi-likelihood approach

```{r}
seedl.quasipois <- glm(Recruits ~ Basal,
                       family = quasipoisson,
                       data = seed)

summary(seedl.quasipois)
```

The dispersion parameter is now a higher value (> 4).

Doesn't actually test this approach becuase resid dev remains unchanged.

## Negative Binomial approximation in GLM

```{r}
library(MASS)

seedl.nb <- glm.nb(Recruits ~ Basal,
                   data = seed)

summary(seedl.nb)
```

You can compare this to the poission using AIC (not quasipossion!)

# Binomial Overdispersion

```{r}
tad <- read.csv("../data/Tadpoles1.csv")
```

```{r}
binom.glm <- glm(cbind(Killed, Initial-Killed) ~ Initial, 
                 family = binomial,
                 data = tad)

summary(binom.glm)
```

## Formal test for overdispersion using chi-sq distribution

```{r}
1 - pchisq(deviance(binom.glm), df.residual(binom.glm))
```

Here, we are using deviance to approximate the chi-squared statistic and calculating the p-value using that and the residual df. This very low p-value signifies that it is overdispersed.

## Quasi likelihood approach

```{r}
qbinom.glm <- glm(cbind(Killed, Initial-Killed) ~ Initial,
                  family = quasibinomial,
                  data = tad)

summary(qbinom.glm)
```

# Beta-binomial example using 'mle2'

GLM doesn't have a beta-binomial, so you can fit it yourself using the MLE approach.

## Set up binomial

```{r}
library(bblme)
library(emdbook)

binomNLL <- function(a, c, N, K){
  -sum(dbinom(K, 
              prob=exp(a*N+C) / (1 + exp(a*N+C)), 
              size=N,
              log=T))
}

binom.fit <- mle2

summary(binomNLL)

```

```{r}
betabinomNLL <- function(a, c, theta, N, K){
  -sum(dbetabinom(K, 
                  size=N,
                  prob=exp(a*N+c) / (1 + exp(a*N+c)),
                  theta=theta,
                  log=T)) #use theta =10
}

betabinom.fit <- mle2(betabinomNLL, start=list(a = -0.008, c = -0.1, theta = 10),
                      data= list(N = tad$Initial, K = tad$Killed))

summary(betabinom.fit)
```


