---
title: "Model Selection"
author: "Michelle Evans"
date: ""
output: html_document
---

```{r setup, include=FALSE}
#wd should be set to Project folder (not RProject)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="../") #set wd to proj directory
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(viridis)
library(nlme)
library(lme4)
library(MuMIn)
# install.packages("R2admb")
# install.packages("glmmADMB", 
#     repos=c("http://glmmadmb.r-forge.r-project.org/repos",
#             getOption("repos")),
#     type="source")
library(glmmADMB)
```

```{r}
emergenceData <- readRDS("data/emergence.RData")
survival <- readRDS("data/survival.RData")
fecundity <- readRDS("data/fecundity.RData")
AedesAll <- readRDS("data/AedesIndividualSurvival.RData")
```

This document goes through model selection for each analysis section of the ms.

# Survival

## Aedes

Survival (or prob. of emergence) is a binary variable per individual mosquito. I use a generalized linear mixed model to explore the effect of Aedes density, Stephensi density, and temperature on emergence, with replicate as a random intercept. The data is subset to that of females (this can differ as males tend to develop more quickly and will emerge first). Predictor variables are normalized/scaled.

Reformat data into successes and failures per jar.
```{r format survival data}
aeSurvival <- emergenceData %>%
  filter(Species == "Aedes" & AeDens!= 0 & Sex == "Female") %>%
  group_by(Replicate, Temp, TempNum, AeDens, StDens, Ratio) %>%
  summarise(success = sum(Number, na.rm = T)) %>%
  mutate(failure = (AeDens/2) - success) %>%
  ungroup() %>%
  #rescale predictor variables
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T))) %>%
  #drop jar that was all male
  filter(!(Replicate == "A" & Temp == "24" & Ratio == "8:24"))
  
#for when more than 50% were female
aeSurvival$failure[aeSurvival$failure<0] <- 0
```

Aedes model selection
```{r}
m0 <- glmer(cbind(success, failure) ~ 1 + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))
m1 <- glmer(cbind(success, failure) ~ AeDensScale + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))
m2 <- glmer(cbind(success, failure) ~ AeDensScale + TempNum + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))
m3 <- glmer(cbind(success, failure) ~ AeDensScale + poly(TempNum,2) + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))
m4 <- glmer(cbind(success, failure) ~ AeDensScale*poly(TempNum,2) + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))

m5 <- glmer(cbind(success, failure) ~ AeDensScale + poly(TempNum,2) + StDensScale + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))

m6 <- glmer(cbind(success, failure) ~ AeDensScale + poly(TempNum,2) * StDensScale + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))

m7 <- glmer(cbind(success, failure) ~ AeDensScale + poly(TempNum,2) + StDensScale + TempNum:StDensScale + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))

m8 <- glmer(cbind(success, failure) ~ AeDensScale*poly(TempNum,2) + StDensScale*poly(TempNum,2) + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))
```

```{r}
modelSummary <- cbind(model = c("m0","m1","m2","m3", "m4", "m5", "m6", "m7" , "m8"), 
                   do.call(rbind, lapply(list(m0, m1, m2, m3, m4, m5, m6, m7, m8), broom::glance)),
                   AICc = AICc(m0, m1, m2, m3, m4, m5, m6, m7, m8), 
                   AICweights = Weights(AIC(m0, m1, m2, m3, m4, m5, m6, m7, m8)))
modelSummary
anova(m6,m7, m8)

AedesSurvivalModel <- m7

```

Based on the above, model 7 (survival ~ AeDens + StDens + Temp^2 + StDens:Temp) fits best. Model 6 fits similary, but has an additional term of an interaction between polynomial temperature and stephensi density, and so the more parismonious model is chosen. Now we explore model residuals.

```{r}
plot(AedesSurvivalModel, id = 0.01, idLabels=~.obs)
qqnorm(resid(AedesSurvivalModel))
qqline(resid(AedesSurvivalModel), col = "red")

aeSurvival$preds <- predict(AedesSurvivalModel)
aeSurvival$resids <- resid(AedesSurvivalModel)

ggplot(data = aeSurvival, aes(x=StDens, y = resids))+
  geom_point() +
  geom_smooth(se=T)+
  theme_base()

ggplot(data = aeSurvival, aes(x=AeDens, y = resids))+
  geom_point() +
  geom_smooth(se=T)+
  theme_base()

ggplot(data = aeSurvival, aes(x=TempNum, y = resids))+
  geom_point() +
  geom_smooth(se=T)+
  theme_base()

```

The residuals for this model are slightly heteroskedastic, but I think are fine in general.

```{r}
summary(AedesSurvivalModel)

confint(AedesSurvivalModel)
```

Plot predicted values in a heatmap.

```{r}
#heatmap plots
newData <- expand.grid(AeDens=seq(0,128, by=2), StDens=seq(0,128, by=2), TempNum=c(16,20,24,28,32), Replicate = c("A", "B"))
newData <- newData %>%
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))

newData$preds <- predict(AedesSurvivalModel, type = "response", newdata = newData)
#get means over replicates to plot
predicted <- newData %>%
  group_by(TempNum, AeDens, StDens) %>%
  summarise(preds = mean(preds))
#remove extrapolation outside of measured range
predicted$preds[predicted$AeDens+predicted$StDens>128] <- NA  

#plot it
ggplot(predicted, aes(x=StDens, y=AeDens, z=preds))+
  geom_raster(aes(fill=preds))+
  geom_contour(color="gray90")+
  theme_minimal()+
  scale_fill_viridis(name="Prop. Emerged")+
  facet_wrap(~TempNum, ncol = 5) +
  xlab("Stephensi Density") +
  ylab("Aegypti Density")
```

## Stephensi

```{r}
#stephensi
stSurvival <- emergenceData %>%
  filter(Species == "Stephensi" & StDens!= 0 & Sex == "Female") %>%
  group_by(Replicate, Temp, TempNum, AeDens, StDens, Ratio) %>%
  summarise(success = sum(Number, na.rm = T)) %>%
  mutate(failure = (StDens/2) - success) %>%
  ungroup() %>%
  #rescale predictor variables
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))
#for when more than 50% were female
stSurvival$failure[stSurvival$failure<0] <- 0
```

Stephensi model selection
```{r}
m0 <- glmer(cbind(success, failure) ~ 1 + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))
m1 <- glmer(cbind(success, failure) ~ StDensScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))
m2 <- glmer(cbind(success, failure) ~ StDensScale + TempScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))
m3 <- glmer(cbind(success, failure) ~ StDensScale + poly(TempScale,2) + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))
m4 <- glmer(cbind(success, failure) ~ StDensScale*poly(TempScale,2) + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))

m5 <- glmer(cbind(success, failure) ~ StDensScale + poly(TempScale,2) + AeDensScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))

m6 <- glmer(cbind(success, failure) ~ StDensScale + poly(TempScale,2) * AeDensScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))

m7 <- glmer(cbind(success, failure) ~ StDensScale + poly(TempScale,2) + AeDensScale + TempScale:StDensScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))

modelSummary <- cbind(model = c("m0","m1","m2","m3", "m4", "m5", "m6", "m7"), 
                   do.call(rbind, lapply(list(m0, m1, m2, m3, m4, m5, m6, m7), broom::glance)),
                   AICc = AICc(m0, m1, m2, m3, m4, m5, m6, m7), 
                   AICweights = Weights(AIC(m0, m1, m2, m3, m4, m5, m6, m7)))
modelSummary

anova(m6,m7)

StephSurvivalModel <- m7
```

For stephensi, model m7 is the best fit (survival ~ AeDens + Temp^2 + StDens + Temp^2*StDens)

```{r}
plot(StephSurvivalModel, id = 0.01, idLabels=~.obs)
qqnorm(resid(StephSurvivalModel))
qqline(resid(StephSurvivalModel), col = "red")

stSurvival$preds <- predict(StephSurvivalModel)
stSurvival$resids <- resid(StephSurvivalModel)

ggplot(data = stSurvival, aes(x=factor(StDens), y = resids))+
  geom_boxplot() +
  geom_line(aes(y=0), color="red") +
  theme_base()

ggplot(data = stSurvival, aes(x=factor(AeDens), y = resids))+
  geom_boxplot() +
  theme_base()

ggplot(data = stSurvival, aes(x=Temp, y = resids))+
  geom_boxplot() +
  theme_base()

ggplot(data = stSurvival, aes(x=Replicate, y = resids))+
  geom_boxplot() +
  theme_base()

```

The residuals on this model look very bad. Not sure what to do.

Now we can predict over the response surface.

```{r}
#heatmap plots
newData <- expand.grid(AeDens=seq(0,128, by=2), StDens=seq(0,128, by=2), TempNum=c(16,20,24,28,32), Replicate = c("A", "B"))
newData <- newData %>%
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))

newData$preds <- predict(StephSurvivalModel, type = "response", newdata = newData)
#get means over replicates to plot
predicted <- newData %>%
  group_by(TempNum, AeDens, StDens) %>%
  summarise(preds = mean(preds))
#remove extrapolation outside of measured range
predicted$preds[predicted$AeDens+predicted$StDens>128] <- NA  

#plot it
ggplot(predicted, aes(x=StDens, y=AeDens, z=preds))+
  geom_raster(aes(fill=preds))+
  geom_contour(color="gray90")+
  theme_minimal()+
  scale_fill_viridis(name="Prop. Emerged")+
  facet_wrap(~TempNum, ncol = 5) +
  xlab("Stephensi Density") +
  ylab("Aegypti Density")
```


# Fecundity

Split into Two Dataframes. Change the NAs of those that didn't lay to 0.

```{r}
aeFec <- dplyr::filter(fecundity, Species == "Aedes")
aeFec$Eggs[is.na(aeFec$Eggs)] <- 0
stFec <- dplyr::filter(fecundity, Species == "Stephensi")
stFec$Eggs[is.na(stFec$Eggs)] <- 0
```


We will use a zero-inflated mixed model here from the `glmmADMB` package because we have many zeros, and this drastically improved model fit during preliminary explorations. Note that these models take a long time to run.

## Aedes

Model Selection

```{r, warning = F}
m0 <- glmmadmb(Eggs ~ 1 + (1|Replicate),
                   data = aeFec,
                    zeroInflation=TRUE, 
                    family="nbinom")

m1 <- glmmadmb(Eggs ~ TempNum + (1|Replicate),
                   data = aeFec,
                    zeroInflation=TRUE, 
                    family="nbinom")

m2 <- glmmadmb(Eggs ~ AeDens + (1|Replicate),
                   data = aeFec,
                    zeroInflation=TRUE, 
                    family="nbinom")

m3 <- glmmadmb(Eggs ~ AeDens + StDens + (1|Replicate),
                   data = aeFec,
                    zeroInflation=TRUE, 
                    family="nbinom")

m4 <- glmmadmb(Eggs ~ AeDens + TempNum + (1|Replicate),
                   data = aeFec,
                    zeroInflation=TRUE, 
                    family="nbinom")

m5 <- glmmadmb(Eggs ~ AeDens + poly(TempNum,2) + (1|Replicate),
                  data = aeFec,
                    zeroInflation=TRUE, 
                    family="nbinom")

m6 <- glmmadmb(Eggs ~ AeDens*poly(TempNum,2) + (1|Replicate),
                   data = aeFec,
                    zeroInflation=TRUE, 
                    family="nbinom")

m7 <- glmmadmb(Eggs ~ AeDens*TempNum + (1|Replicate),
                   data = aeFec,
                    zeroInflation=TRUE, 
                    family="nbinom")

```

```{r}
modelSummary <- cbind(model = c("m0","m1","m2","m3", "m4", "m5", "m6", "m7"), 
                   logLik = logLik(m0, m1, m2, m3, m4, m5, m6, m7),
                   AIC = AIC(m0, m1, m2, m3, m4, m5, m6, m7),
                   AICc = AICc(m0, m1, m2, m3, m4, m5, m6, m7), 
                   AICweights = Weights(AIC(m0, m1, m2, m3, m4, m5, m6, m7)))
modelSummary
AedesFecundityModel <- m5

```

The best fitting model is m5 (AeDens + TempNum^2).

```{r}
summary(AedesFecundityModel)

confint(AedesFecundityModel)
```


```{r}
#plot(AedesFecundityModel, id = 0.01, idLabels=~.obs)
plot(fitted(AedesFecundityModel), residuals(AedesFecundityModel))
# qqnorm(resid(AedesFecundityModel))
# qqline(resid(AedesFecundityModel), col = "red")

aeFec$preds <- predict(AedesFecundityModel)
aeFec$resids <- resid(AedesFecundityModel)

ggplot(data = aeFec, aes(x = StDens, y = resids, group=StDens))+
  geom_boxplot() +
  theme_base()

ggplot(data = aeFec, aes(x = AeDens, y = resids, group = AeDens))+
  geom_boxplot() +
  theme_base()

ggplot(data = aeFec, aes(x = TempNum, y = resids, group = TempNum))+
  geom_boxplot() +
  theme_base()
```

Create heatmaps of the eggs. If there are only two covariates, can do a 3D surface plot.

```{r}
#heatmap plots
newData <- expand.grid(AeDens=seq(0,128, by=2), StDens=seq(0,128, by=2), TempNum=c(16,20,24,28,32), Replicate = c("A", "B"))
newData <- newData %>%
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))

newData$preds <- predict(AedesFecundityModel, type = "response", newdata = newData)
#get means over replicates to plot
predicted <- newData %>%
  group_by(TempNum, AeDens, StDens) %>%
  summarise(preds = mean(preds))
#remove extrapolation outside of measured range
predicted$preds[predicted$AeDens+predicted$StDens>128] <- NA  

#plot it
ggplot(predicted, aes(x=StDens, y=AeDens, z=preds))+
  geom_raster(aes(fill=preds))+
  geom_contour(color="gray90")+
  theme_minimal()+
  scale_fill_viridis(name="Number of Eggs Laid")+
  facet_wrap(~TempNum, ncol = 5) +
  xlab("Stephensi Density") +
  ylab("Aegypti Density")
```

## Stephensi

There are too many zeros in this for the zero-inflated model to work. 

```{r}
stFec <- dplyr::filter(fecundity, Species == "Stephensi" & !is.na(Eggs))
```


```{r, include = F}
ggplot(data = stFec, aes(x = TempNum, y = Eggs))+
  geom_point()
```

Model Selection

```{r}
m0 <- glmer(Eggs ~ 1 + (1|Replicate),
                   data = stFec,
                   family = gaussian)

m1 <- glmer(Eggs ~ poly(TempNum,2) + (1|Replicate),
                   data = stFec,
                   family = gaussian)

m2 <- glmer(Eggs ~ AeDens + (1|Replicate),
                   data = stFec,
                   family = gaussian)

m3 <- glmer(Eggs ~ StDens + (1|Replicate),
                   data = stFec,
                   family = gaussian)

m4 <- glmer(Eggs ~ AeDens + TempNum + (1|Replicate),
                   data = stFec,
                   family = gaussian)

m5 <- glmer(Eggs ~ AeDens + poly(TempNum,2) + (1|Replicate),
                   data = stFec,
                   family = gaussian)

m6 <- glmer(Eggs ~ AeDens*poly(TempNum,2) + (1|Replicate),
                   data = stFec,
                   family = gaussian)

m7 <- glmer(Eggs ~ (AeDens + StDens)*TempNum + (1|Replicate),
                   data = stFec,
                   family = gaussian)
```

```{r}
modelSummary <- cbind(model = c("m0","m1","m2","m3", "m4", "m5", "m6", "m7"), 
                   do.call(rbind, lapply(list(m0, m1, m2, m3, m4, m5, m6, m7), broom::glance)),
                   AICc = AICc(m0, m1, m2, m3, m4, m5, m6, m7), 
                   AICweights = Weights(AIC(m0, m1, m2, m3, m4, m5, m6, m7)))
modelSummary

StephensiFecundityModel <- m6
```

The best fitting model is (Aedes*Temp^2), however none of these covariates are significant (i.e. their confidence intervals overlap 0). So Stephensi fecundity is not affected by inter or intraspecific densities or temperature, but note that this may also be because of the small sample size.

```{r}
summary(StephensiFecundityModel)
confint(StephensiFecundityModel)
```

```{r}
plot(StephensiFecundityModel, id = 0.01, idLabels=~.obs)
qqnorm(resid(StephensiFecundityModel))
qqline(resid(StephensiFecundityModel), col = "red")

stFec$preds <- predict(StephensiFecundityModel)
stFec$resids <- resid(StephensiFecundityModel)

hist(stFec$resids)

ggplot(data = stFec, aes(x = StDens, y = resids, group = StDens))+
  geom_boxplot() +
  theme_base()

ggplot(data = stFec, aes(x = AeDens, y = resids, group = AeDens))+
  geom_boxplot() +
  theme_base()

ggplot(data = stFec, aes(x=TempNum, y = resids, group = TempNum))+
  geom_boxplot() +
  theme_base()
```

Create heatmaps of the eggs:

```{r, include = F}
#heatmap plots
newData <- expand.grid(AeDens=seq(0,128, by=2), StDens=seq(0,128, by=2), TempNum=c(16,20,24,28,32), Replicate = c("A", "B"))
newData <- newData %>%
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))

newData$preds <- predict(m0, type = "response", newdata = newData)
#get means over replicates to plot
predicted <- newData %>%
  group_by(TempNum, AeDens, StDens) %>%
  summarise(preds = mean(preds))
#remove extrapolation outside of measured range
predicted$preds[predicted$AeDens+predicted$StDens>128] <- NA  

#plot it
ggplot(predicted, aes(x=StDens, y=AeDens, z=preds))+
  geom_raster(aes(fill=preds))+
  geom_contour(color="gray90")+
  theme_minimal()+
  scale_fill_viridis(name="Number of Eggs Laid")+
  facet_wrap(~TempNum, ncol = 5) +
  xlab("Stephensi Density") +
  ylab("Aegypti Density")
```

# Growth Rate

Growth rate calculated via Sugihara based on raw data.

## Aedes

Calculate growth rate. Using raw data except for fecundity (since this was only estimated from a subset). Fecundity is estimated using models above. In the case of stephensi, it is just the mean across all treatments, since none of our covariates were significant.

```{r}
# only need female emergence
GrowthAedes <- emergenceData %>%
  filter(Species =="Aedes" & Sex == "Female") %>%
  #assume 50% female
  mutate(N0 = AeDens/2) %>%
  rename(x = Day, Ax = Number) %>%
  #don't calculate for ones that had no aedes to begin with
  filter(AeDens>0) 

#add in predicted fecundity
GrowthAedes$fwx <- predict(AedesFecundityModel, type = "response", newdata = GrowthAedes)

#calculate overall per capita growth rate
GrowthAedes <- GrowthAedes %>%
  mutate(Axfwx = Ax*fwx, xAxfwx = x*Ax*fwx) %>%
  group_by(Replicate, Temp, Ratio, TempNum, AeDens, StDens) %>%
  mutate_at(c("Axfwx", "xAxfwx"), sum, na.rm=T) %>%
  mutate(D=14) %>%
  ungroup() %>%
  mutate(r=(log((1/N0)*Axfwx))/(D+(xAxfwx/Axfwx))) %>%
  group_by(Replicate, Temp, TempNum, Ratio, AeDens, StDens, Species) %>%
  #each row is now a duplicate of the same thing so only need the first one
  slice(1) %>%
  mutate(lambda=exp(r)) %>%
  ungroup() %>%
    #rescale predictor variables
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T))) %>%
  #drop jar that was all male
  filter(!(Replicate == "A" & Temp == "24" & Ratio == "8:24")) %>%
  #drop unneccesary columns
  dplyr::select(-x, -Ax, - Sex, - N0, -fwx, - Axfwx, - xAxfwx, -D)
```

```{r}
ggplot(data =  GrowthAedes, aes(x = AeDens, y = lambda, color = Temp)) +
  geom_point() +
  geom_smooth()

ggplot(data =  GrowthAedes, aes(x = Temp, y = lambda, color = AeDens)) +
  geom_point() +
  geom_smooth()

hist((GrowthAedes$lambda))
```

Model Selection

```{r}
#model selection
m0 <- glmer(lambda ~ 1 + (1|Replicate),
                   data = GrowthAedes,
                   family = gaussian)

m1 <- glmer(lambda ~ poly(TempScale,2) + (1|Replicate),
                   data = GrowthAedes,
                   family = gaussian)

m2 <- glmer(lambda ~ AeDensScale + (1|Replicate),
                   data = GrowthAedes,
                   family = gaussian)

m3 <- glmer(lambda ~ poly(AeDensScale,2)+ (1|Replicate),
                   data = GrowthAedes,
                   family = gaussian)

m4 <- glmer(lambda ~ AeDensScale + poly(TempScale,2) + (1|Replicate),
                   data = GrowthAedes,
                   family = gaussian)

m5 <- glmer(lambda ~ AeDensScale + poly(TempScale,2) + StDensScale + (1|Replicate),
                   data = GrowthAedes,
                   family = gaussian)

m6 <- glmer(lambda ~ AeDensScale*TempScale+ (1|Replicate),
                   data = GrowthAedes,
                   family = gaussian)

m7 <- glmer(lambda ~ poly(AeDensScale,2) + poly(TempScale,2) + (1|Replicate),
                   data = GrowthAedes,
                   family = gaussian)

```

```{r}
modelSummary <- cbind(model = c("m0","m1","m2","m3", "m4", "m5", "m6", "m7"), 
                   do.call(rbind, lapply(list(m0, m1, m2, m3, m4, m5, m6, m7), broom::glance)),
                   AICc = AICc(m0, m1, m2, m3, m4, m5, m6, m7), 
                   AICweights = Weights(AIC(m0, m1, m2, m3, m4, m5, m6, m7)))
modelSummary
anova(m4, m7, m8)

AedesGrowthModel <- m7
```

The final model is Aedes^2 + Temp^2.

```{r}
summary(AedesGrowthModel)
confint(AedesGrowthModel)
```

```{r}
plot(AedesGrowthModel, id = 0.01, idLabels=~.obs)
qqnorm(resid(AedesGrowthModel))
qqline(resid(AedesGrowthModel), col = "red")

GrowthAedes$preds <- predict(AedesGrowthModel)
GrowthAedes$resids <- resid(AedesGrowthModel)

hist(GrowthAedes$resids)

ggplot(data = GrowthAedes, aes(x=AeDens, y = resids, group = AeDens))+
  geom_boxplot() +
  theme_base()

ggplot(data = GrowthAedes, aes(x=TempNum, y = resids, group = TempNum))+
  geom_boxplot() +
  theme_base()

```

Create heatmaps of the growth. Ditto 3D surface above:

```{r}
#heatmap plots
newData <- expand.grid(AeDens=seq(0,128, by=2), StDens=seq(0,128, by=2), TempNum=c(16,20,24,28,32), Replicate = c("A", "B"))
newData <- newData %>%
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))

newData$preds <- predict(AedesGrowthModel, type = "response", newdata = newData)
#get means over replicates to plot
predicted <- newData %>%
  group_by(TempNum, AeDens, StDens) %>%
  summarise(preds = mean(preds))
#remove extrapolation outside of measured range
predicted$preds[predicted$AeDens+predicted$StDens>128] <- NA  

#plot it
ggplot(predicted, aes(x=StDens, y=AeDens, z=preds))+
  geom_raster(aes(fill=preds))+
  geom_contour(color="gray90")+
  theme_minimal()+
  scale_fill_viridis(name="Growth Rate")+
  facet_wrap(~TempNum, ncol = 5) +
  xlab("Stephensi Density") +
  ylab("Aegypti Density") 
```

## Stephensi

```{r}
# only need female emergence
GrowthStephensi <- emergenceData %>%
  filter(Species =="Stephensi" & Sex == "Female") %>%
  #assume 50% female
  mutate(N0 = StDens/2) %>%
  rename(x = Day, Ax = Number)  %>%
  filter(StDens>0)

#add in predicted fecundity based on overall mean since model had no significance
GrowthStephensi$fwx <- mean(fecundity$Eggs[fecundity$Species == "Stephensi"], 
                            na.rm = T)

#calculate overall per capita growth rate
GrowthStephensi <- GrowthStephensi %>%
  mutate(Axfwx = Ax*fwx, xAxfwx = x*Ax*fwx) %>%
  group_by(Replicate, Temp, Ratio, TempNum, AeDens, StDens) %>%
  mutate_at(c("Axfwx", "xAxfwx"), sum, na.rm=T) %>%
  mutate(D=14) %>%
  ungroup() %>%
  mutate(r=(log((1/N0)*Axfwx))/(D+(xAxfwx/Axfwx))) %>%
  group_by(Replicate, Temp, TempNum, Ratio, AeDens, StDens, Species) %>%
  #each row is now a duplicate of the same thing so only need the first one
  slice(1) %>%
  mutate(lambda=exp(r)) %>%
  ungroup() %>%
    #rescale predictor variables
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T))) %>%
  #drop unneccesary columns
  dplyr::select(-x, -Ax, - Sex, - N0, -fwx, - Axfwx, - xAxfwx, -D)

# NAs are truly zeros
GrowthStephensi$lambda[is.na(GrowthStephensi$lambda)] <- 0  
```

```{r}
ggplot(data =  GrowthStephensi, aes(x = StDens, y = lambda, color = Temp)) +
  geom_point() +
  geom_smooth(se = F)

ggplot(data =  GrowthStephensi, aes(x = Temp, y = lambda, color = StDens)) +
  geom_point() +
  geom_smooth(se = F)

ggplot(data =  GrowthStephensi[GrowthStephensi$lambda>0,], 
       aes(x = StDens, y = lambda, color = Temp)) +
  geom_point()

ggplot(data =  GrowthStephensi[GrowthStephensi$lambda>0,], 
       aes(x = Temp, y = lambda, color = StDens)) +
  geom_point()

```

Model Selection

Impliment a [hurdle model](http://seananderson.ca/2014/05/18/gamma-hurdle.html).

```{r}
#create nonzero variable
GrowthStephensi$nonZero <- ifelse(GrowthStephensi$lambda>0,1,0)

plot(GrowthStephensi$StDens, GrowthStephensi$nonZero, pch=20)
abline(lm(nonZero ~ StDens, data = GrowthStephensi))

plot(GrowthStephensi$TempNum, GrowthStephensi$nonZero, pch=20)
abline(lm(nonZero ~ TempNum, data = GrowthStephensi)) #temperature has no effect zero

hist(GrowthStephensi$lambda)
hist(GrowthStephensi$lambda[GrowthStephensi$lambda>0])
```

Model selection for zero vs. nonzero growth
```{r}
m0 <- glmer(nonZero ~ 1 + (1|Replicate),
            data = GrowthStephensi,
            family = binomial(link = logit))

m1 <- glmer(nonZero ~ TempScale + (1|Replicate),
            data = GrowthStephensi,
            family = binomial(link = logit))

m2 <- glmer(nonZero ~ AeDensScale + (1|Replicate),
            data = GrowthStephensi,
            family = binomial(link = logit))

m3 <- glmer(nonZero ~ StDensScale + (1|Replicate),
            data = GrowthStephensi,
            family = binomial(link = logit))

m4 <- glmer(nonZero ~ AeDensScale + StDensScale + (1|Replicate),
            data = GrowthStephensi,
            family = binomial(link =logit))

m5 <- glmer(nonZero ~ AeDensScale * StDensScale + (1|Replicate),
            data = GrowthStephensi,
            family = binomial(link = logit))
```

```{r}
modelSummary <- cbind(model = c("m0","m1","m2","m3", "m4", "m5"), 
                   do.call(rbind, lapply(list(m0, m1, m2, m3, m4, m5), broom::glance)),
                   AICc = AICc(m0, m1, m2, m3, m4, m5), 
                   AICweights = Weights(AIC(m0, m1, m2, m3, m4, m5)))
modelSummary

StephensiGrowthModelBin <- m4
```

Whether or not growth is zero (basically mosquitoes emerge), is dependent on density (Aedes and Stephensi), not temperature.

Inspect residuals, although this is more important for final hurdle model.
```{r}
plot(StephensiGrowthModelBin, id = 0.01, idLabels=~.obs)
qqnorm(resid(StephensiGrowthModelBin))
qqline(resid(StephensiGrowthModelBin), col = "red")

GrowthStephensi$predsBin <- predict(StephensiGrowthModelBin)
GrowthStephensi$residsBin <- resid(StephensiGrowthModelBin)

hist(GrowthStephensi$residsBin)

ggplot(data = GrowthStephensi, aes(x=AeDens, y = residsBin, group = AeDens))+
  geom_boxplot() +
  theme_base()

ggplot(data = GrowthStephensi, aes(x=TempNum, y = residsBin, group = TempNum))+
  geom_boxplot() +
  theme_base()
```

Second part of hurdle model for non-zeros
```{r}
m0 <- glmer(lambda ~ 1 + (1|Replicate),
            data = subset(GrowthStephensi, nonZero==1),
            family = gaussian)

m1 <- glmer(lambda ~ TempScale + (1|Replicate),
            data = subset(GrowthStephensi, nonZero==1),
            family = gaussian)

m2 <- glmer(lambda ~ AeDensScale + (1|Replicate),
            data = subset(GrowthStephensi, nonZero==1),
            family = gaussian)

m3 <- glmer(lambda ~ StDensScale + (1|Replicate),
            data = subset(GrowthStephensi, nonZero==1),
            family = gaussian)

m4 <- glmer(lambda ~ StDensScale + TempScale + (1|Replicate),
            data = subset(GrowthStephensi, nonZero==1),
            family = gaussian)

m5 <-  glmer(lambda ~ StDensScale * TempScale + (1|Replicate),
            data = subset(GrowthStephensi, nonZero==1),
            family = gaussian)

m6 <- glmer(lambda ~ StDensScale + poly(TempScale,2) + (1|Replicate),
            data = subset(GrowthStephensi, nonZero==1),
            family = gaussian)
```

```{r}
modelSummary <- cbind(model = c("m0","m1","m2","m3", "m4", "m5", "m6"), 
                   do.call(rbind, lapply(list(m0, m1, m2, m3, m4, m5, m6), broom::glance)),
                   AICc = AICc(m0, m1, m2, m3, m4, m5, m6), 
                   AICweights = Weights(AIC(m0, m1, m2, m3, m4, m5, m6)))
modelSummary

StephensiGrowthModel2 <- m1
```

Best Model is Temp + StDens

```{r}
plot(StephensiGrowthModel2, id = 0.01, idLabels=~.obs)
qqnorm(resid(StephensiGrowthModel2))
qqline(resid(StephensiGrowthModel2), col = "red")

GrowthStephensi$predsGam[GrowthStephensi$nonZero==1] <- predict(StephensiGrowthModel2)
GrowthStephensi$residsGam[GrowthStephensi$nonZero==1] <- resid(StephensiGrowthModel2)

hist(GrowthStephensi$residsGam)

ggplot(data = GrowthStephensi, aes(x=AeDens, y = residsGam, group = AeDens))+
  geom_boxplot() +
  theme_base()

ggplot(data = GrowthStephensi, aes(x=TempNum, y = residsGam, group = TempNum))+
  geom_boxplot() +
  theme_base()
```


```{r}
summary(StephensiGrowthModel2)
confint(StephensiGrowthModel2)
```


```{r}
unHurdle <- function(mod1, mod2, newDF){
  #' Function that creates predictions based on binomial/guassian hurdle model
  
  #' @param mod1 binomial model
  #' @param mod2 second model, gaussian in this case
  #' @param newDF new data frame to predict over, must have same columns as original
  phi_zero <- predict(mod1, newdata = newDF, type = "response")
  mod2Preds <- predict(mod2, newdata = newDF, type = "response")
  #calculate probability of zero in second model to subtract from first
  phi_count <- pnorm(q = 0, 
                    mean = mod2Preds,
                    sd = sigma(mod2),
                    log.p = F)
  phi <- phi_zero - phi_count
  predsAll <- phi * mod2Preds
  return(predsAll)
}

predsAll <- unHurdle(mod1 = StephensiGrowthModelBin, mod2 = StephensiGrowthModel2, newDF = GrowthStephensi)

GrowthStephensi$predsFinal <- predsAll
```

```{r}
#heatmap plots
newData <- expand.grid(AeDens=seq(0,128, by=2), StDens=seq(0,128, by=2), TempNum=c(16,20,24,28,32), Replicate = c("A", "B"))
newData <- newData %>%
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))

newData$preds <- unHurdle(mod1 = StephensiGrowthModelBin, mod2 = StephensiGrowthModel2, newDF = newData)
#get means over replicates to plot
predicted <- newData %>%
  group_by(TempNum, AeDens, StDens) %>%
  summarise(preds = mean(preds))
#remove extrapolation outside of measured range
predicted$preds[predicted$AeDens+predicted$StDens>128] <- NA  

#plot it
ggplot(predicted, aes(x=StDens, y=AeDens, z=preds))+
  geom_raster(aes(fill=preds))+
  geom_contour(color="gray90")+
  theme_minimal()+
  scale_fill_viridis(name="Growth Rate")+
  facet_wrap(~TempNum, ncol = 5) +
  xlab("Stephensi Density") +
  ylab("Aegypti Density") 
```

