---
title: "Model Selection"
author: "Michelle Evans"
date: "11/22/2017"
output: html_document
---

```{r setup, include=FALSE}
#wd should be set to Project folder (not RProject)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="../") #set wd to proj directory
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(viridis)
library(nlme)
library(lme4)
library(MuMIn)
```

```{r}
emergenceData <- readRDS("data/emergence.RData")
survival <- readRDS("data/survival.RData")
fecundity <- readRDS("data/fecundity.RData")
AedesAll <- readRDS("data/AedesIndividualSurvival.RData")
```

This document goes through model selection for each analysis section of the ms.

# Survival

## Aedes

Survival (or prob. of emergence) is a binary variable per individual mosquito. I use a generalized linear mixed model to explore the effect of Aedes density, Stephensi density, and temperature on emergence, with replicate as a random intercept. The data is subset to that of females (this can differ as males tend to develop more quickly and will emerge first). Predictor variables are scaled.

Reformat data into successes and failures per jar.
```{r format survival data}
aeSurvival <- emergenceData %>%
  filter(Species == "Aedes" & AeDens!= 0 & Sex == "Female") %>%
  group_by(Replicate, Temp, TempNum, AeDens, StDens, Ratio) %>%
  summarise(success = sum(Number, na.rm = T)) %>%
  mutate(failure = (AeDens/2) - success) %>%
  ungroup() %>%
  #rescale predictor variables
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T))) %>%
  #drop jar that was all male
  filter(!(Replicate == "A" & Temp == "24" & Ratio == "8:24"))
  
#for when more than 50% were female
aeSurvival$failure[aeSurvival$failure<0] <- 0
```

Aedes model selection
```{r}
m0 <- glmer(cbind(success, failure) ~ 1 + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))
m1 <- glmer(cbind(success, failure) ~ AeDensScale + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))
m2 <- glmer(cbind(success, failure) ~ AeDensScale + TempScale + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))
m3 <- glmer(cbind(success, failure) ~ AeDensScale + poly(TempScale,2) + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))
m4 <- glmer(cbind(success, failure) ~ AeDensScale*poly(TempScale,2) + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))

m5 <- glmer(cbind(success, failure) ~ AeDensScale + poly(TempScale,2) + StDensScale + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))

m6 <- glmer(cbind(success, failure) ~ AeDensScale + poly(TempScale,2) * StDensScale + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))

m7 <- glmer(cbind(success, failure) ~ AeDensScale + poly(TempScale,2) + StDensScale + TempScale:StDensScale + (1|Replicate),
                   data = aeSurvival,
                   family = binomial(link = "logit"))

modelSummary <- cbind(model = c("m0","m1","m2","m3", "m4", "m5", "m6", "m7"), 
                   do.call(rbind, lapply(list(m0, m1, m2, m3, m4, m5, m6, m7), broom::glance)),
                   AICc = AICc(m0, m1, m2, m3, m4, m5, m6, m7), 
                   AICweights = Weights(AIC(m0, m1, m2, m3, m4, m5, m6, m7)))
modelSummary
anova(m6,m7)

AedesSurvivalModel <- m7

```

Based on the above, model 7 (survival ~ AeDens + StDens + Temp^2 + StDens:Temp) fits best. Model 6 fits similary, but has an additional term of an interaction between polynomial temperature and stephensi density, and so the more parismonious model is chosen. Now we explore model residuals.

```{r}
plot(AedesSurvivalModel, id = 0.01, idLabels=~.obs)
qqnorm(resid(AedesSurvivalModel))
qqline(resid(AedesSurvivalModel), col = "red")

aeSurvival$preds <- predict(AedesSurvivalModel)
aeSurvival$resids <- resid(AedesSurvivalModel)

ggplot(data = aeSurvival, aes(x=StDens, y = resids))+
  geom_point() +
  geom_smooth(se=T)+
  theme_base()

ggplot(data = aeSurvival, aes(x=AeDens, y = resids))+
  geom_point() +
  geom_smooth(se=T)+
  theme_base()

ggplot(data = aeSurvival, aes(x=TempNum, y = resids))+
  geom_point() +
  geom_smooth(se=T)+
  theme_base()

```

The residuals for this model are slightly heteroskedastic, but I think are fine in general.

```{r}
summary(AedesSurvivalModel)

confint(AedesSurvivalModel)
```

Plot predicted values in a heatmap.

```{r}
#heatmap plots
newData <- expand.grid(AeDens=seq(0,128, by=2), StDens=seq(0,128, by=2), TempNum=c(16,20,24,28,32), Replicate = c("A", "B"))
newData <- newData %>%
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))

newData$preds <- predict(AedesSurvivalModel, type = "response", newdata = newData)
#get means over replicates to plot
predicted <- newData %>%
  group_by(TempNum, AeDens, StDens) %>%
  summarise(preds = mean(preds))
#remove extrapolation outside of measured range
predicted$preds[predicted$AeDens+predicted$StDens>128] <- NA  

#plot it
ggplot(predicted, aes(x=StDens, y=AeDens, z=preds))+
  geom_raster(aes(fill=preds))+
  geom_contour(color="gray90")+
  theme_minimal()+
  scale_fill_viridis(name="Prop. Emerged")+
  facet_wrap(~TempNum, ncol = 5) +
  xlab("Stephensi Density") +
  ylab("Aegypti Density")
```

## Stephensi

```{r}
#stephensi
stSurvival <- emergenceData %>%
  filter(Species == "Stephensi" & StDens!= 0 & Sex == "Female") %>%
  group_by(Replicate, Temp, TempNum, AeDens, StDens, Ratio) %>%
  summarise(success = sum(Number, na.rm = T)) %>%
  mutate(failure = (StDens/2) - success) %>%
  ungroup() %>%
  #rescale predictor variables
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))
#for when more than 50% were female
stSurvival$failure[stSurvival$failure<0] <- 0
```

Stephensi model selection
```{r}
m0 <- glmer(cbind(success, failure) ~ 1 + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))
m1 <- glmer(cbind(success, failure) ~ StDensScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))
m2 <- glmer(cbind(success, failure) ~ StDensScale + TempScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))
m3 <- glmer(cbind(success, failure) ~ StDensScale + poly(TempScale,2) + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))
m4 <- glmer(cbind(success, failure) ~ StDensScale*poly(TempScale,2) + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))

m5 <- glmer(cbind(success, failure) ~ StDensScale + poly(TempScale,2) + AeDensScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))

m6 <- glmer(cbind(success, failure) ~ StDensScale + poly(TempScale,2) * AeDensScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))

m7 <- glmer(cbind(success, failure) ~ StDensScale + poly(TempScale,2) + AeDensScale + TempScale:StDensScale + (1|Replicate),
                   data = stSurvival,
                   family = binomial(link = "logit"))

modelSummary <- cbind(model = c("m0","m1","m2","m3", "m4", "m5", "m6", "m7"), 
                   do.call(rbind, lapply(list(m0, m1, m2, m3, m4, m5, m6, m7), broom::glance)),
                   AICc = AICc(m0, m1, m2, m3, m4, m5, m6, m7), 
                   AICweights = Weights(AIC(m0, m1, m2, m3, m4, m5, m6, m7)))
modelSummary

anova(m6,m7)

StephSurvivalModel <- m7
```

For stephensi, model m6 is the best fit (survival ~ AeDens + Temp^2*StDens)

```{r}
plot(StephSurvivalModel, id = 0.01, idLabels=~.obs)
qqnorm(resid(StephSurvivalModel))
qqline(resid(StephSurvivalModel), col = "red")

stSurvival$preds <- predict(StephSurvivalModel)
stSurvival$resids <- resid(StephSurvivalModel)

ggplot(data = stSurvival, aes(x=factor(StDens), y = resids))+
  geom_boxplot() +
  geom_line(aes(y=0), color="red") +
  theme_base()

ggplot(data = stSurvival, aes(x=factor(AeDens), y = resids))+
  geom_boxplot() +
  theme_base()

ggplot(data = stSurvival, aes(x=Temp, y = resids))+
  geom_boxplot() +
  theme_base()

ggplot(data = stSurvival, aes(x=Replicate, y = resids))+
  geom_boxplot() +
  theme_base()

```

The residuals on this model look very bad. Not sure what to do.

Now we can predict over the response surface.

```{r}
#heatmap plots
newData <- expand.grid(AeDens=seq(0,128, by=2), StDens=seq(0,128, by=2), TempNum=c(16,20,24,28,32), Replicate = c("A", "B"))
newData <- newData %>%
  mutate(TempScale = as.vector(scale(TempNum, center = F, scale = T))) %>%
  mutate(AeDensScale = as.vector(scale(AeDens, center = F, scale = T))) %>%
  mutate(StDensScale = as.vector(scale(StDens, center = F, scale = T)))

newData$preds <- predict(StephSurvivalModel, type = "response", newdata = newData)
#get means over replicates to plot
predicted <- newData %>%
  group_by(TempNum, AeDens, StDens) %>%
  summarise(preds = mean(preds))
#remove extrapolation outside of measured range
predicted$preds[predicted$AeDens+predicted$StDens>128] <- NA  

#plot it
ggplot(predicted, aes(x=StDens, y=AeDens, z=preds))+
  geom_raster(aes(fill=preds))+
  geom_contour(color="gray90")+
  theme_minimal()+
  scale_fill_viridis(name="Prop. Emerged")+
  facet_wrap(~TempNum, ncol = 5) +
  xlab("Stephensi Density") +
  ylab("Aegypti Density")
```


# Fecundity

# Growth Rate


