---
title: "Assignment 4"
author: "Michelle Evans"
date: "October 30, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#packages
library(dplyr)
library(ggplot2)
library(lme4)
library(nlme)
```

```{r lme overdispersion function}
overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```


# 1

Create function from script. Note that Plotfact is now included in the model under column PlotID:

```{r}
createDF <- function(N=32){
  splist <- c('A', 'B', 'C')
  Plot <- sample(1:5, N, replace=TRUE)
  tree <- data.frame(Plot)
  tree$Species <- factor(sample(splist, N, replace = TRUE))
  tree$Y <- 0 # Placeholder for Y (growth rate)
  alpha.A <- 8 # species A effect
  alpha.B <- 6 # species B effect
  alpha.C <- 4.5 # species C effect
  sigma.sq <- 2 # standard deviation of epsilon error term
  sigma.sq.b <- 0.5
  b <- rnorm(5, mean=0, sd=sigma.sq.b) #plot level effects
  epsilon <- rnorm(N, mean=0, sd=sigma.sq) # generate error terms
  for (k in 1:N){
    if (tree$Species[k]=="A") alpha <- alpha.A
    else if (tree$Species[k]=="B") alpha <- alpha.B
    else alpha <- alpha.C
    tree$Y[k] <- alpha + b[tree$Plot[k]] + epsilon[k] # Growth model
  }
  tree$PlotID <- factor(tree$Plot) # Tell R that Plot is a categorical variable
  return(tree)
}
```

Create simulated data (N=32)

```{r}
set.seed(8675309)
tree <- createDF(N=32)
```

Visualize data:

```{r}
hist(tree$Y)

ggplot(data=tree, aes(x=Species, y=Y))+
  geom_boxplot()+
  facet_wrap(~PlotID)
```

Analyze using `lmer` model:

```{r}
mod0 <- lmer(Y~1 + (1|PlotID),
             data=tree)

mod1 <- lmer(Y~Species + (1|PlotID),
            data=tree)
summary(mod1)

anova(mod0, mod1)
plot(mod1)
overdisp_fun(mod1)
confint(mod1)
```

The model is overdispersed, however the effect of Species is significant, as measured via a comparison to the null model with an anova and 95% confidence intervals.

# 2

Aggregate data
```{r}
treePlot <- tree %>%
  group_by(PlotID, Species) %>%
  summarise(growth = mean(Y))
```

Re-analyze:
```{r}
modNew0 <- lmer(growth ~ 1 + (1|PlotID),
               data = treePlot)
modNew1 <- lmer(growth ~ Species + (1|PlotID),
               data = treePlot)
summary(modNew1)

anova(modNew0, modNew1)
plot(modNew1)
overdisp_fun(modNew1)
confint(modNew1)
```

The results are qualitatively similar to those above. The effect of species is significant in both cases, with Species A > Species B > Species C in growth rates. However, the model at the plot level is not overdispersed as the one at the level of the individual tree is.

# 3

```{r}
set.seed(8675309)
Nvec <- c(32, 64, 100, 150, 200)
results <- data.frame(N = Nvec, 
                      pvalueInd = NA, 
                      pvaluePlot = NA)
for (i in 1:length(Nvec)){
  treeDF <- createDF(N = Nvec[i])
  
  modInd <- lme(Y ~ Species,
                random = ~1|PlotID,
                data = treeDF)
  
  treePlot <- treeDF %>%
    group_by(Species, PlotID)%>%
    summarise(Y = mean(Y))
  
  modPlot <- lme(Y ~ Species,
                 random = ~1|PlotID,
                 data = treePlot)
  
  results$pvalueInd[i] <- anova(modInd)["Species", "p-value"]
  results$pvaluePlot[i] <- anova(modPlot)["Species", "p-value"]
}

plot(results$N, results$pvaluePlot, col = NA, xlab = "N", ylab = "p-value")
lines(results$N, results$pvaluePlot, col = "dodgerblue")
lines(results$N, results$pvalueInd, col = "forestgreen")
legend(col = c("forestgreen", "dodgerblue"), 
       lty = c(1,1), 
       legend = c("Individual Level", "Plot Level"), 
       "topright")
mtext("lme results for different values of N")
```

Analyzing data at the plot level reduces the power of the analysis because information is lost by aggregating to the mean, however as N increases, this becomes less of an issue because the variance within a species becomes smaller (even though information is being lost, the mean is more precise). It is important to note, however, that this is the result of one simulation, and rerunning it with different seeds results in some qualitatively different results.