---
title: "Data Import and Management"
output: html_document
date: "2017-08-16"
---

These is the excercises from ECOL8990 on 2017-08-16. All exercises (or most) have also been repeated using the `tidyverse`. These dataframes are denoted with a T appended to the end.

```{r}
library(reshape2)
library(tidyr)
library(dplyr)
library(stringr)
```

# Intro

```{r}
sm <- read.csv("data/Ketter_soil_moisture_2014.csv")

sm.long <- melt(sm, id.vars=c("Day", "Bin"), variable.name= "Depth", value.name= "GWC")
#using tidyverse
sm.longTidy <- sm %>%
  gather(key=Depth, value=GWC, GWC_30cm, GWC_10cm)

sm.long <- sm.long[order(sm.long$Bin, sm.long$Day, sm.long$Depth),]

#tidyverse
sm.longT <- sm.long %>%
  select(Bin, Day, Depth, GWC) %>%
  mutate(newDepth=case_when(
    Depth == "GWC_30cm" ~ 30,
    Depth == "GWC_10cm" ~ 10
  )) %>%
  select(-Depth) %>%
  rename(Depth=newDepth)

#write.csv(sm.longT, "data/Ketter_sm2014_long.csv", row.names=F)

```

```{r}
rm(list=ls())
```

Read in file from before

```{r}
sm.long <- read.csv("data/Ketter_sm2014_long.csv")
```

```{r}
sm.wide <- dcast(sm.long, Bin+Day~Depth, value.var="GWC")

names(sm.wide)[3:4] <- c("GWC_10cm", "GWC_30cm")
sm.wide$Delta_GWC <- sm.wide$GWC_10cm - sm.wide$GWC_30cm

#  Using tidyverse
sm.wideT <- sm.long %>%
  spread(key=Depth, value=GWC) %>%
  rename(GWC_10cm=`10`, GWC_30cm=`30`) %>%
  mutate(Delta_GWC=GWC_10cm-GWC_30cm)
```

# Exercise 1

```{r}
dataFiles <- list.files("data/", full.names = T)
corv2006 <- read.csv(dataFiles[1])
corv2007 <- read.csv(dataFiles[2])

corv2006Long <- corv2006 %>%
  gather(key=species, value=abundance, Sp1:Sp5) %>%
  mutate(year=2006)

corv2007Long <- corv2007 %>%
  gather(key=species, value=abundance, Sp1:Sp7) %>%
  mutate(year=2007)

corvAll <- rbind(corv2006Long, corv2007Long)
```

#Data Logger Data

```{r}
nyl <- read.csv("data/CR1000_SM_5_min.dat", skip=3)
colnames(nyl) <- c("Timestamp", "RN", "VWC1_5", "VWC1_10","VWC1_20", "VWC1_40", "VWC1_70", "VWC2_10", "VWC3_10", "VWC4_10", "Precip")

time <- strptime(nyl$Timestamp, "%Y-%m-%d %H:%M:%S")
objects(time)
nyl$Day <- time$yday

nyl.sm <- nyl %>%
  group_by(Day) %>%
  summarise_at(mean, .vars=vars(-Timestamp, -Precip, -RN)) 

nyl.Precip <- nyl %>%
  group_by(Day) %>%
  summarise(PrecipSum=sum(Precip))

nyl.ag <- full_join(nyl.sm, nyl.Precip, by="Day")
```

# Exercise 2

Turn nyl.sm into long format
```{r}
nyl.sm.l <- nyl.sm %>%
  gather(key=ProbeID, value=PercVWC, VWC1_5:VWC4_10)
class(nyl.sm.l$Day)
class(nyl.sm.l$ProbeID)
```

Split location and depth
```{r}
nyl.sm.split <- nyl.sm.l %>%
  separate(ProbeID, c("Loc", "Depth"), sep="_", remove=F) %>%
  select(Day, ProbeID, VWC=PercVWC, Loc, Depth)
head(nyl.sm.split)
nyl.sm.split$Depth <- as.numeric(nyl.sm.split$Depth)
```

Pair with GPS Locations
```{r}
gps <- read.csv("data/Nylsvlei_Probe_GPS.csv")
```

Subset 10 cm locations
```{r}
nyl.sm.10 <- nyl.sm.split %>%
  filter(Depth==10) %>%
  group_by(Loc) %>%
  summarise(meanVWC=mean(VWC))
```

Merge with GPS data
```{r}
nyl.sm.merge <- full_join(gps, nyl.sm.10, by="Loc")
```

Plot (base)
```{r}
radius <- sqrt(nyl.sm.merge$meanVWC/pi)
symbols(nyl.sm.merge$X, nyl.sm.merge$Y, circles=radius, inches=0.25, xlab="Eastings (m)", ylab="Northings (m)")
```

Plot (ggplot)
```{r}
ggplot(data=nyl.sm.merge, aes(x=X, y=Y))+
  geom_point(aes(size=sqrt(meanVWC)/pi), fill=NA)+
  xlab("Eastings (m)")+
  ylab("Nothings (m)")+
  theme_base()+
  guides(size=F)
```

# Extra exercise

Soil moisture as a function of depth, location, and precipitation (three panel)
