---
title: "Final Project"
author: "Michelle Evans"
date: ""
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = F, warning = F, message = F)
library(ggplot2)
library(ggthemes)
```

```{r set ggplot theme}
theme_set(theme_bw())
```


# Introduction

The frequency of infectious disease emergence has been increasing since the mid-20th century. Incidence of vector borne diseases, specifically, have increased over 25% over the past decade [@jones2008], with notable recent outbreaks such as Zika and chikungunya in the Americas. One driver of this increase is land-use change, such as agricultural intensification and urbanization [@gottdenker2014, @jones2013]. Urbanization can cause increases in vector-borne disease through changes to mosquito habitat and environmental conditions. For container breeding species such as *Aedes aegypti*, the primary vector of dengue and yellow fever, urbanization increases the amount of habitat available for breeding [@zahouli2017, @li2014]. Further, urbanization can alter the microclimate of the environment through the urban heat island effect, leader to warmer temperatures in urban areas[@peng2012]. 

Mosquitoes, being small ectotherms, are extremely susceptible to changes in temperature, which can significantly impact on life history traits relevant to demography and disease transmission [@murdock2017, @paaijmans2009]. The majority of past work has focused on the abiotic effects of urbanization, ignoring changes to biotic factors such as community composition and larval densities.  Mosquito population growth is density dependent in the larval stage [@juliano1998, reiskind2009a], and can be mediated by environmental factors such as temperature [@costanzo2005, lounibos2002a] , leading to a potential interaction between abiotic and biotic factors.

Many primary vectors of arboviruses are container breeding species, which often cohabit in containers with other mosquito species, leading to strong control via density-dependence and interspecific competition [@juliano2010]. For example, the invasion of the Southeastern United States by *Ae. albopictus* has lead to habitat partitioning with *Ae. aeygypti*, due to *Ae. albopictus*'s competitive dominance at low resource levels [@fader2016]. In this case, the vectors share many pathogens, and the dominance of an area by one species is not likely to introduce novel arboviruses to the population. However, in some systems, cohabiting vectors may transmit different pathogens, and competitive outcomes can have significant impacts on disease transmission. In urban India, for example, *Anopheles stephensi*, the primary vector of urban malaria, is often found cohabiting with *Ae. aegypti*, the primary vector of dengue [@thomas2016]. SENTENCE ABOUT CYCLICAL NATURE OF DISEASE (Olivier Telle paper, maybe something else). This seasonality of disease has been hypothesized to be tied to outcomes of competition, however, the outcome of competition between *An. stephensi* and *Ae. aeygypti*, and the extent to which it is seasonally dependent is unknown.

Here, we explore the interactive effects of interspecific competition, intraspecific competition, and temperature on the population growth of two co-habiting mosquito species, *Aedes aegypti* and *Anopheles stephensi*, using a laboratory model system to parameterize demographic life history traits and ultimately calculate population grwoth rates. SOMETHING SOMETHING the relative importance of abiotic and biotic factors on mosquito population dynamics.

# Materials/Methods

## Mosquito Rearing

We used a response surface design to rear  *Aedes aegypti* (Mexico, F5) and *Anopheles stephensi* (Walter Reed strain) first-instar larvae across 5 temperatures at 15 densities for a total of 75 treatments. There were three total density levels (32, 64, and 128 individuals), each with 5 ratios of Aedes:Stephensi (0:4,1:3,2:2,3:1,4:0) (FIGURE). The experimental setup was replicated twice.

```{r fig1, fig.width=3, fig.height=3, fig.cap="Plot of density treatments used in experiment. Each group of fifteen density treatments was repeated across five temperatures (16C, 20C, 24C, 28C, 32C)."}
ggplot(data=data.frame(Aedes=c(0,8,16,24,32,0,16,32,48,64,0,32,64,96,128), 
                       Stephensi = c(32,24,16,8,0,64,48,32,16,0,128,96,64,32,0)))+
  geom_point(aes(x = Stephensi, y = Aedes), size=6)+
  ylab("Aedes Density")+
  xlab("Stephensi Density")
```


Larvae were reared in 32 oz. glass jars with 250mL deionized water and 0.10g of pellet fish food (Hikari Cichlid Gold, baby size). Jars were covered with a fine mesh and placed in Percival incubators with a daily periodic fluctuation of 9&deg;C following the Parton-Logan equation, characterized by a sine wave during the daytime and an exponential curve during the nighttime [@parton1981]. Incubators were set to 80\% relative humidity and 12:12 hour light:dark cycle.

To record information on mosquito survival and development rates, the daily numbers, species, and sex of mosquitoes were recorded per species ratio and temperature treatment.

To record measures of adult female fecundity and longevity, we offered a subset of emerging mosquitoes a blood meal and followed individuals for the span of their life. The subset encompassed mosquitoes emerging during the period of peak emergence to ensure an adequate sample size and a similarly aged cohort at the time of the blood feed (4-6 days old). Adults were stored in reach-in Percival incubators at a constant 27&deg;C, 80\% relative humidity, 12:12 hr light:dark cycle and offered a 10\%  sucrose solution *ad libitum*. 48 hours prior to the blood meal, the sucrose solution was replaced with deionized water, which was removed 24 hours prior to the blood meal to encourage higher feeding rates. Blood meals consisting of whole human blood were administered through a water-jacketed membrance feeder kept at 38&deg;C. 

Mosquitoes were allowed to feed for 20 minutes, after which blood fed females were sorted into individual 50 mL plastic centrifuge tubes. Moistened cotton was placed at the bottom of each tube, and covered with a filter paper to collect eggs. Each tube was covered with a fine mesh and kept in a walk-in incubator at the same environmental conditions noted above. Individual females were monitored daily for mortality and egg-laying events, after which the eggs were counted and the moistened cotton and filter paper removed. Females continued to be kept as above and offered a 10\% sucrose solution until death.

## Calculating Population Growth Rates

We calculated the per capita population growth rate (Equation \ref{eq:1}) per species ratio and temperature following [@livdahl1984]:

\begin{equation} \label{eq:1}
r' = \frac{ln(\frac{1}{N_0}\sum_{x}^{ }{A_x}f(\bar{w_x}))}{D+\frac{\sum_{x}^{ }xA_xf(\bar{w_x})}{\sum_{x}^{ }A_xf(\bar{w_x})}}
\end{equation}

Where $N_0$ is the initial number of female mosquitoes (assumed to be 50\% of the larvae), $A_x$ is the number of mosquitoes emerging on day $x$, $D$ is the time to reproduction following emergence, and $f(\bar{w_x})$ is the mean fecundity as measured by egg counts.

## Fitting Models and Elasticity Analyses



# Results

```{r, eval = F, include = F}
load("finalModels.RData")
#create tidy table of results

```


# Discussion

# Acknowledgements

# References